Cell Decompositions for Rank Two Quiver Grassmannians

### 2018-03-13 Grassmannian_cells.tex
Fixed some typos in proof of Corollary 5.18

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Combinatorial description of non-empty cells}
\label{sec:combinatorics}
\noindent In this section, we provide two combinatorial descriptions of the non-empty cells in the quiver Grassmannians of (truncated) preprojective representations of $K(n)$.
The first is quiver theoretic and follows directly from the recursive construction of the cell decomposition in Section~\ref{sec:fibrations}.
The second is the notion of compatible pairs in a maximal Dyck path arising in the computation of rank 2 cluster variables \cite{llz}.
We give a bijection between these which provides a partial geometric explanation for the combinatorial construction of counting polynomials for rank two quiver Grassmannians given in \cite{rupel}.

@@

\begin{corollary}
  \label{cor:piecewise not compatible}
  For $m\ge3$ and $0\le r\le n-1$, consider $S_H\subset H_{m+1}^{[r]}$ and $S_V\subset V_{m+1}^{[r]}$ so that $(S_H,S_V)$ is piecewise compatible.
  Assume $\big(S_H^{[r+1]},S_V^{[r+1]}\big)\in\cC_{m+1}^{[r+1]}$.
  Then $(S_H,S_V)$ is not compatible if and only if $H_{m,r+1,n}^{[r+1]}\subset S_{H,r+1}$ and $V_{m+1}^{[r+1]}\subset S_V$.
\end{corollary}
\begin{proof}
  We begin with the reverse implication.
  First note that there are $(n-r)u_m-u_{m-1}$ horizontal edges and $(n-r)u_{m-1}-u_{m-2}$ vertical edges in $D_{m+1}^{[r]}$. 
  It follows that $H_{m,r+1,n}^{[r+1]}\sqcup H_{m+1}^{[r+1]}$ contains $n(n-r)u_{m-1}-nu_{m-2}$ horizontal edges and $V_{m-1}^{[r+1]}\sqcup V_{m+1}^{[r+1]}$ contains $n(n-r)u_{m-2}-nu_{m-3}$ vertical edges (note that $H_{m,r+1,n}^{[r+1]}$ naturally identifies with the Dyck path $H_{m-1}^{[r+1]}$).

@@

  To continue, let $(S_H,S_V)$ be a pair for $D_{m+1}$, $m\ge4$, which is not compatible.
  Write $\varphi:H_m\to V_{m+1}$ for the bijection given by $\varphi(h_i)=v_i$ for $1\le i\le u_m$.
  For any subset $T\subset H_m$, set $\varphi^*(T)=V_{m+1}\setminus\varphi(T)$.
  Clearly, the map $\varphi^*$ gives a bijection between subsets of $H_m$ and subsets of $V_{m+1}$.
  In Section 3.2 of \cite{rupel}, a new pair of subsets $\big((\varphi^*)^{-1}S_V,\Omega^{-1}S_H\big)$ for $D_m$ is given, we refer the reader to \emph{loc. cit} for notation.
  By \cite[Proposition 3.10]{rupel}, the pair $\big((\varphi^*)^{-1}S_V,\Omega^{-1}S_H\big)$ is not compatible, but is piecewise compatible by \cite[proposition 3.16]{rupel}.
  Thus by induction, we must have $H_{m-2}^{[r+1]}\subset(\varphi^*)^{-1}S_V$ and $V_m^{[r+1]}\subset\Omega^{-1}S_H$.
  It follows from piecewise compatibility that $H_m^{[r+1]}\cap(\varphi^*)^{-1}S_V=\varnothing$.
  But then by the definition of $\varphi^*$ we have $V_{m-1}^{[r+1]}\cap S_V=\varnothing$ and $S_V^{[r+1]}=V_{m+1}^{[r+1]}$ so that $D(v_{u_m};S_V)=hv_{u_m}$ with $h$ as in the first case above.
  Since $(S_H,S_V)$ is not compatible, Theorem~\ref{th:blocking edge conditions} states that $h$ must be the blocking edge for $S_H$ and we must have $D(h;S_H)=hv_{u_m}$.
  But this can only occur if $H_{m,r+1,n}^{[r+1]}\subset S_H$ since $H_{m+1}^{[r+1]}\cap S_H=\varnothing$ by piecewise compatibility.
\end{proof}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Combinatorial Descriptions of Non-Empty Cells}
\label{sec:combinatorics}
\noindent In this section, we provide two combinatorial descriptions of the non-empty cells in the quiver Grassmannians of (truncated) preprojective representations of $K(n)$.
The first is quiver theoretic and follows directly from the recursive construction of the cell decomposition in Section~\ref{sec:fibrations}.
The second is the notion of compatible pairs in a maximal Dyck path arising in the computation of rank 2 cluster variables \cite{llz}.
We give a bijection between these which provides a partial geometric explanation for the combinatorial construction of counting polynomials for rank two quiver Grassmannians given in \cite{rupel}.

@@

\begin{corollary}
  \label{cor:piecewise not compatible}
  For $m\ge3$ and $0\le r\le n-1$, consider $S_H\subset H_{m+1}^{[r]}$ and $S_V\subset V_{m+1}^{[r]}$ so that $(S_H,S_V)$ is piecewise compatible.
  Assume $\big(S_H^{[r+1]},S_V^{[r+1]}\big)\in\cC_{m+1}^{[r+1]}$.
  Then $(S_H,S_V)$ is not compatible if and only if $H_{m,r+1,n}^{[r+1]}\subset S_{H,r+1}$ and $V_{m+1}^{[r+1]}\subset S_V$.
\end{corollary}
\begin{proof}
  We begin with the reverse implication.
  First note that there are $(n-r)u_m-u_{m-1}$ horizontal edges and $(n-r)u_{m-1}-u_{m-2}$ vertical edges in $D_{m+1}^{[r]}$. 
  It follows that $H_{m,r+1,n}^{[r+1]}\sqcup H_{m+1}^{[r+1]}$ contains $n(n-r)u_{m-1}-nu_{m-2}$ horizontal edges and $V_{m-1}^{[r+1]}\sqcup V_{m+1}^{[r+1]}$ contains $n(n-r)u_{m-2}-nu_{m-3}$ vertical edges (note that $D_{m,r+1,n}^{[r+1]}$ naturally identifies with the Dyck path $D_{m-1}^{[r+1]}$).

@@

  To continue, let $(S_H,S_V)$ be a pair for $D_{m+1}$, $m\ge4$, which is not compatible.
  Write $\varphi:H_m\to V_{m+1}$ for the bijection given by $\varphi(h_i)=v_i$ for $1\le i\le u_m$.
  For any subset $T\subset H_m$, set $\varphi^*(T)=V_{m+1}\setminus\varphi(T)$.
  Clearly, the map $\varphi^*$ gives a bijection between subsets of $H_m$ and subsets of $V_{m+1}$.
  In Section 3.2 of \cite{rupel}, a new pair of subsets $\big((\varphi^*)^{-1}S_V,\Omega^{-1}S_H\big)$ for $D_m$ is given, we refer the reader to \emph{loc.\ cit} for notation.
  By \cite[Proposition 3.10]{rupel}, the pair $\big((\varphi^*)^{-1}S_V,\Omega^{-1}S_H\big)$ is not compatible, but is piecewise compatible by \cite[Proposition 3.16]{rupel}.
  Thus by induction, we must have $H_{m-1,r+1,n}^{[r+1]}\subset(\varphi^*)^{-1}S_V$ and $V_m^{[r+1]}\subset\Omega^{-1}S_H$.
  It follows from piecewise compatibility that $H_m^{[r+1]}\cap(\varphi^*)^{-1}S_V=\varnothing$.
  But then by the definition of $\varphi^*$ we have $V_{m,r+1,n}^{[r+1]}\cap S_V=\varnothing$ and $S_V^{[r+1]}=V_{m+1}^{[r+1]}$ so that $D(v_{u_m};S_V)=hv_{u_m}$ with $h$ as in the first case above.
  Since $(S_H,S_V)$ is not compatible, Theorem~\ref{th:blocking edge conditions} states that $h$ must be the blocking edge for $S_H$ and we must have $D(h;S_H)=hv_{u_m}$.
  But this can only occur if $H_{m,r+1,n}^{[r+1]}\subset S_H$ since $H_{m+1}^{[r+1]}\cap S_H=\varnothing$ by piecewise compatibility.
\end{proof}
